\begin{abstract}
% Fine-grained semantic segmentation is inherently challenging, as it combines the difficulties of precise localization with subtle inter-class distinctions. This challenge is further exacerbated by the highly imbalanced class distribution of the FungiTastic dataset. A straightforward approach is to leverage SAM3 by prompting it with specific fungal classes; however, this method suffers from limited scalability, slow inference, and low IoU due to frequent class omissions. Moreover, its computational cost grows with both the number of images and the number of classes.

% To address these limitations, we propose a two-stage framework. First, SAM3 is used to generate class-agnostic segments. Then, a foundation model (DINOv3) extracts latent representations for each segment, enabling classification in embedding space. This decoupling of segmentation and classification improves scalability and flexibility. We demonstrate that our approach is effective across one-shot, few-shot, and full-data regimes, maintaining strong performance across varying levels of supervision.

% \todo{Furthermore, we show that we can leverage the segments to perform augmentations on the few-shot examples and combine them to further boost performances! -> IF TRUE}

% The code will be publicly released upon acceptance.
Fine-grained semantic segmentation is inherently challenging, as it requires both precise localization and the ability to distinguish subtle inter-class differences. This challenge is further amplified by two factors: 1) the highly imbalanced class distribution of the FungiTastic dataset and 2) the highly nuisance variation due to wildly different image acquisition settings. 

A straightforward solution is to prompt SAM3 with class-specific queries; however, this approach suffers from limited scalability, slow inference, and low IoU due to frequent class omissions. Additionally, its computational cost increases with both the number of images and the number of classes.

To address these limitations, we propose a parameter-free two-stage framework that decouples segmentation from classification. In the first stage, SAM3 is employed to generate class-agnostic segmentations through macro taxonomy queries. In the second stage, a foundation model (DINOv3) extracts latent representations for each segment, enabling classification in embedding space. 

We then devise a parameter-free approach which leverages class prototypes coupled with a correction on the intrinsic geometry of the DINOv3 embedding space. This design removes dependence on class-specific prompts (which are not available during inference) improves scalability and unlock lightweight flexibility thanks to the parameter free property.

We then report baseline results across low data regimes spanning from one-shot, few-shot to few hundred images per class. We think it is important to ground the work on real world scenarios where data regime for segmentation is the de-facto standard due to the costly annotation process. To the best of our knowledge we think this work constitutes the first baseline on such problem.

\end{abstract}