\section{Results}
\label{sec:results}

\paragraph{Result flow.}
After the teaser comparisons, the full README-defined paper is supposed to answer three follow-up questions: which frozen backbone family is strongest, whether diversity matters more than quantity when several backbones are fused, and which lightweight classifier extracts the most from those features. Those broader ablations are still pending, so the concrete evidence in this draft comes from the completed DINOv3 ViT-7B slice below rather than from placeholder plots.

\begin{table*}[t]
    \centering
    \small
    \setlength{\tabcolsep}{7pt}
    \begin{tabular}{lccc}
        \toprule
        Feature variant & 224 test acc. (\%) & 448 test acc. (\%) & Note \\
        \midrule
        \texttt{cls} & 72.22 & 74.30 & global token baseline \\
        \texttt{patch} & 64.40 & 65.68 & whole-image patch mean \\
        \texttt{masked} & 72.05 & 72.12 & oracle mask pooled patches \\
        \texttt{cls+patch} & 71.97 & 74.01 & global + whole-image pooled patches \\
        \texttt{cls+masked} & \textbf{74.04} & \textbf{75.53} & global + oracle masked patches \\
        \texttt{register} & 67.56 & -- & exploratory 224-only register token \\
        \texttt{cls+register} & 71.74 & -- & exploratory 224-only fusion \\
        \bottomrule
    \end{tabular}
    \caption{Completed DINOv3 ViT-7B MLP slice currently available in the repo. The main README-aligned comparison is between global, pooled, masked, and fused descriptors; the latest 224 rerun also logs register-based variants.}
    \label{tab:mlp-results}
\end{table*}

\paragraph{Completed segmentation slice: DINOv3 + MLP.}
\Cref{tab:mlp-results} retains the completed slice that already exists in the repo. The main pattern is consistent across both resolutions: \texttt{cls+masked} is the best configuration, outperforming plain \texttt{cls} by 1.82 points at 224 and 1.23 points at 448. Mask-aware pooling alone stays close to \texttt{cls} at 224 but falls behind at 448, while plain patch averaging is clearly weak. In the latest 224 rerun, the exploratory register variants also trail \texttt{cls+masked}, so the strongest current evidence still favors global context plus segmented local evidence.

\begin{table}[t]
    \centering
    \small
    \setlength{\tabcolsep}{8pt}
    \begin{tabular}{lcc}
        \toprule
        Modality & Best $k$ & Test acc. (\%) \\
        \midrule
        \texttt{cls} & 9 & 52.66 \\
        \texttt{patch} & 10 & 33.50 \\
        \texttt{masked} & 9 & 50.98 \\
        \texttt{cls+patch} & 10 & 49.46 \\
        \texttt{cls+masked} & 10 & \textbf{52.86} \\
        \bottomrule
    \end{tabular}
    \caption{Best 224 $k$-NN accuracy over $k=1,\dots,10$ for the completed DINOv3 slice.}
    \label{tab:knn-results}
\end{table}

\paragraph{Completed classifier slice: DINOv3 + 224 $k$-NN.}
The current classifier comparison is only partially populated, but the 224 $k$-NN sweep already preserves the same rough ranking. \Cref{tab:knn-results} shows that \texttt{cls+masked} again gives the best result, narrowly exceeding \texttt{cls}, while \texttt{patch} is far behind. This is useful because it suggests that the segmentation effect is not unique to the MLP head.

\paragraph{What is still missing from the final story.}
The README-level paper is larger than the completed slice. The 200-class oracle teaser is now exported, but the 2000-class SAM-3 teaser, the cross-family backbone sweep, the fusion ablation, and the prototype and linear-probe baselines are still missing. The current paper therefore reports a real first result slice while keeping those missing pieces explicit.
