\section{Conclusion and Future Work}
We presented a simple yet effective training-free baseline for fine-grained semantic segmentation in low-data regimes, built on a two-stage decomposition of the problem into class-agnostic segmentation and fine-grained classification. By combining SAM3 macro-taxonomy masks with DINOv3 prototype-based recognition, our approach removes the need for class-specific prompting at inference time and yields a significantly more scalable alternative to exhaustive fine-grained querying. On FungiTastic, this design provides a practical baseline for a particularly challenging setting characterized by subtle inter-class differences, long-tailed class distributions, and strong nuisance variability across acquisition conditions.

Our results show that the main bottleneck is not only the scarcity of supervision, but also the geometry of the embedding space used for prototype matching. Off-the-shelf DINOv3 [CLS] features are not sufficiently aligned with the metric structure required for fine-grained discrimination in this setting. In contrast, correcting the feature geometry through PCA whitening leads to a marked improvement, as it suppresses nuisance-dominated high-variance directions and allows more class-informative components to contribute to cosine-based comparison. This finding suggests that, in low-data fine-grained regimes, representation preprocessing can be as important as the choice of the underlying foundation model itself.

Beyond the specific benchmark, our study highlights a broader message: strong fine-grained dense prediction baselines can emerge from carefully composed foundation models, even without any task-specific training. We hope this work provides a useful reference point for future research on low-data fine-grained segmentation. While FungiTastic provides a challenging setting due to subtle inter-class differences and high variability, its current segmentation setup is still limited to a single fungal class per image. Future work will therefore aim to further strengthen the fine-grained capability of the proposed pipeline by complementing global [CLS] descriptors with [PATCH] tokens masked by the general-class SAM3 segmentations, evaluating alternative visual backbones, and testing the method on additional benchmarks.

Overall, we believe this work constitutes a first baseline for fine-grained semantic segmentation under realistic low-data conditions, and highlights the potential of carefully composed training-free pipelines when their design is matched to the structure of the problem.